---
title: "Data_607-Assignment_1"
author: "Nnaemezue Obi-Eyisi"
date: "February 2, 2017"
output: html_document
---

Let's first load the Mushroom Dataset into R Dataframe

```{r}
data_url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data"

mushroom_df <- read.table(file=data_url, header = FALSE, sep =",",
                          stringsAsFactors = FALSE)

#View(mushroom_df)
```

Based on the rules in the data dictionary we care about
Attribute odor V6
Attribute spore-print-color V21
Attribute stalk-surface-below-ring V14
Attribute stalk-color-above-ring V13
Attribute habitat V23
Attribute cap-color V4
Attribute population V22

Therefore, these attributes will form our subset


Create Mushroom Dataframe subset(Mushroom_df_sub) and name the columns
```{r}
mushroom_df_sub <-mushroom_df[c(1,6,4,15,14,21,22,23)]
names(mushroom_df_sub)<- c("Mushroom_class","odor","cap-color","stalk-color-above-ring"
                           ,"stalk-surface-below-ring","spore-print-color","population",
                           "habitat")

#View top 10 rows of the subset of Mushroom_df
head(mushroom_df_sub, n=10)
```

Transform Mushroom class field to a more descriptive format

However let's see distinct values of the mushroom_class and know that we are transforming all the possible values
```{r}
unique(mushroom_df_sub$Mushroom_class)
mushroom_df_sub$Mushroom_class[c(mushroom_df_sub$Mushroom_class =='p')] <- c("poisonous")
mushroom_df_sub$Mushroom_class[c(mushroom_df_sub$Mushroom_class =='e')] <- c("edible")


```
Transform odor field to a more descriptive format
```{r}
unique(mushroom_df_sub$odor)

lookup_odor<-c(a="almond",l="anise",c="creosote",y="fishy",f="foul", m="musty",n="none",p="pungent",s="spicy")

#Notice if we list the names of the vector we can repeat the colnames
lookup_odor[c('a','l','l')]

#Create a character vector from the column names of the dataframe
head(lookup_odor[mushroom_df_sub$odor],n=10)

#Replace the odor column  with the character vector created but with no names

mushroom_df_sub$odor <- unname(lookup_odor[mushroom_df_sub$odor])



```
Transform cap_color column
```{r}
unique(mushroom_df_sub$`cap-color`) #Ensure all field values have description

lookup_cap.color <- c(n="brown",b="buff",c="cinnamon",g="gray",r="green", p="pink", u="purple",e="red",w="white",y="yellow" )

mushroom_df_sub$`cap-color` <- unname(lookup_cap.color[mushroom_df_sub$`cap-color`])


```
Transform rest of columns
```{r}
lookup_cap.sporeprintcolor <- c(k="black",n="brown",b="buff",h="chocolate",r="green", o="orange",
                      u="purple",w="white",y="yellow" )

mushroom_df_sub$`spore-print-color` <- unname(lookup_cap.sporeprintcolor[mushroom_df_sub$`spore-print-color`])

unique(mushroom_df_sub$`stalk-color-above-ring`)
lookup_stalk_c_abR <- c(n='brown',b='buff',c='cinnamon',g='gray',o='orange',
                                  p='pink',e='red',w='white',y='yellow')
mushroom_df_sub$`stalk-color-above-ring` <- unname(lookup_stalk_c_abR[mushroom_df_sub$`stalk-color-above-ring`])


lookup_stalk_sbr <- c(f='fibrous',y='scaly',k='silky',s='smooth')
mushroom_df_sub$`stalk-surface-below-ring` <- unname(lookup_stalk_sbr[mushroom_df_sub$`stalk-surface-below-ring`])

lookup_population <- c(a='abundant',c='clustered',n='numerous',
                                  s='scattered',v='several',y='solitary')
mushroom_df_sub$population <- unname(lookup_population[mushroom_df_sub$population])

lookup_habitat <- c(g='grasses',l='leaves',m='meadows',p='paths',
                                  u='urban',w='waste',d='woods')
mushroom_df_sub$habitat <- unname(lookup_habitat[mushroom_df_sub$habitat])

```

Display final result of transformed Data frame top 100
```{r}
 head(mushroom_df_sub,20)
```