---
title: "Assignment 5 Data607"
author: "Nnaemezue Obi-Eyisi"
date: "March 1, 2017"
output: html_document
---

#Load libraries for tidyr and dplyr
```{r results='hide', message=FALSE, warning = FALSE}

library("tidyr")
library(dplyr)
library(ggplot2)


```

#Read info from csv into R
```{r readin, eval = TRUE, warning = FALSE}
filename <- c("https://raw.githubusercontent.com/nobieyi00/CUNY_MSDA_R/master/flight_dataset.csv")

flight_dataset_df <- read.csv(filename,stringsAsFactors = FALSE)
flight_dataset_df
```
#Tidy data

#Narrow down the table
```{r}
Flight_tidy <- tbl_df(flight_dataset_df)

#Narrow down the table structure

flight_freq <- Flight_tidy %>% 
  gather (State,frequency,-Airline,-Arrival_Status)

flight_freq
```
#Normalize flight data set so that we can see for each airline and state, the frequencies of ontime and delays 
```{r}
on_time_ds<- select (
        filter(flight_freq, Arrival_Status=='On time' )
        ,Airline,State,On_time_freq=frequency
       )

delayed_ds <-select (
        filter(flight_freq, Arrival_Status=='Delayed' )
        ,Airline,State,Delayed_freq=frequency
       )
flight_ds <-inner_join(on_time_ds, delayed_ds, by = c("Airline","State"))
flight_ds
```

#Mutate Result set to calculate Ratio of delays per Airline, per state

```{r}
Ratio_delay_ds <- mutate(flight_ds, Total_freq= On_time_freq + Delayed_freq, 
        Average_Delays = Delayed_freq/Total_freq)
```

#Rearrange / sort dataset by the delay ratio to see Airline that performs worse

```{r}
arrange(Ratio_delay_ds,State,Average_Delays)  
```

#We start noticing that the AMWEST airline performs worse per state.

```{r}

ggplot(Ratio_delay_ds,aes(x = State, y =Average_Delays))+
      geom_point(aes(color=Airline),               # colour depends on cond2
               size=3) 

```

```{r}

Flight_graph <-mutate(Ratio_delay_ds, Airline_State= paste(Airline, State, sep ="_")) 



ggplot(Flight_graph, aes(x=State,y=Average_Delays)) + geom_boxplot(aes(color=Airline))

```

#In summary, we can say that the ALASKA airline is more on time

#Find total sum of average of the delay for both airline
```{r}
Flight_summary <- Ratio_delay_ds %>% 
  group_by(Airline) %>% 
  summarise(Total_average_Delay = sum(Average_Delays))

Flight_summary

```


```{r}
ggplot(Flight_summary, aes(y=Total_average_Delay,x=Airline))  + geom_boxplot(aes(color= Airline),size =3)
```

#Most delays comes from AMWEST airline

#Just for curiosity, which city experienced the most delays?

```{r}
City_summary <- Ratio_delay_ds %>% 
  group_by(State) %>% 
  summarise(Total_average_Delay = sum(Average_Delays))

arrange(City_summary,Total_average_Delay)
```

#Represent in a box plot
```{r}
ggplot(City_summary, aes(y=Total_average_Delay,x=State))  + geom_boxplot(aes(color= State),size =3)
```

#We can see it is San Franciso
