---
title: "Assigment_2_Data607"
author: "Nnaemezue Obi-Eyisi"
date: "February 12, 2017"
output: html_document
---

Load Required Libraries or packages for Code

```{r}

library(RODBC)
require(ggplot2)

```

Load the SQL table into an R Dataframe
To Set up SQL Server Database ODBC connection
Follow the below steps as prescribed in https://technet.microsoft.com/en-us/library/cc879308(v=sql.105).aspx

To run the SQL Server DSN Configuration Wizard
Open the ODBC Data Source Administrator as described in the topic, How to: Open the ODBC Data Source Administrator.
On the User DSN tab or the System DSN tab, click Add to add a user data source or a system data source.
Click SQL Native Client or SQL Server Native Client 10.0 to select the driver, and then click Finish.
Follow the instructions in the Wizard.
```{r}
SQL_Server_connection <- odbcConnect("SQL_MEZUE-PC")
# SELECT * FROM Movie Reviews table into a Data Frame

# Check that connection is working 
odbcGetInfo(SQL_Server_connection)

# Query the database and put the results into the data frame "dataframe"
 Movie_Review_Df <- sqlQuery(SQL_Server_connection, "
 SELECT *
 FROM
 MyDb.[dbo].[Movie_Reviews_Stage]")
 
 Movie_Review_Df

 
```

 
 Normalized version of the data frame is below
 
```{r}

 
 Reviewer_df <- sqlQuery(SQL_Server_connection, "
 SELECT *
 FROM
 MyDb.[dbo].[Reviewer]")
 
 Reviewer_df
 
 Movies_Title_df <- sqlQuery(SQL_Server_connection, "
 SELECT *
 FROM
 MyDb.[dbo].[Movies]")
 
Movies_Title_df

 Movie_Review_Xref_df <- sqlQuery(SQL_Server_connection, "
 SELECT *
 FROM
 MyDb.[dbo].[Movie_Review_Xref]")

 Movie_Review_Xref_df
```


Perform few data analysis on the Average Ratings from Reviews




```{r}
hist(Movie_Review_Df$Ratings,main ="Movie_Review", xlab = "Ratings")

```



Perform few data analysis on the Average Ratings from Reviews




```{r}
ggplot(data = Movie_Review_Df) +geom_histogram(aes(x=Ratings))
```



Display box plot of the Ratings per Title



```{r}
boxplot(Ratings~ Movie_Title,  data = Movie_Review_Df)
```



Plot points of Movie Ratings




```{r}
p1 <- ggplot(Movie_Review_Df, aes(x = Movie_Title, y = Ratings))

p1 + geom_point(color="red") 
```



Violin Plot of the Average Ratings per movie




```{r}
ggplot(Movie_Review_Df, aes(x=Movie_Title,y=Ratings)) + geom_violin() + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
```
